Bounds on the dimension of consistent convex surrogates

### 2020-05-26 neurips-20/cvx-flats.tex
Checking in for the night.  Will take another pass in the morning

\begin{definition}[Flat]
	A \emph{flat} $F$ of $\reals^n$ is an affine subspace of $\reals^n$.
	In other words, $F\subseteq \reals^n$ can be written $F=\{x\in\reals^n : Wx + b = 0\}$ for some $W\in\reals^{d\times n}$ and $b\in\reals^d$ where $d\leq n$.
	The dimension of the flat $F$ is given by $f := n - \mathrm{rank}(W)$.
\end{definition}

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
	For all $r\in\gamma(p)$, if there is no $(n - d-1)$-dimensional flat $F$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}[Proof]
	We prove the contrapositive: if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $(n-d-1)$-dimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

@@

	Observe that the Minkowski sum $\oplus_y p_y \partial L(u,y) = \{p_1 x_1 + \ldots + p_n x_n : x_1 \in \partial L(u,1), \ldots, x_n \in \partial L(u,n) \}$.
	In order for $\vec 0$ to be in this $p$-weighted Minkowski sum, we then have that each choice of $x_i$ is a flat for $\partial L(u, i)$.
	Moreover, with $p$ fixed, we know there is a vector of witnesses $W' = [x_1, \ldots, x_n] \in \reals^{d\times n}$ such that $\sum_y p_y x_y = \vec 0$.

	Observe that $F' := \ker(W')$ is at most $(n-d)$ dimensions in $\reals^n$ and contains $p$ by construction.
	Since we are only concerned with a flat intersected with the simplex being a subset of $\gamma_r$, we can reduce the dimension by adding an affine simplex constraint. 
	That is, take $F := \{x \in \reals^n : W x + b = \vec 0\}$, where we have $W := [W'; \mathbbm{1}^T] \in \reals^{(d+1) \times n}$ and $b \in \reals^{d+1}$.
	The dimension of $F$ is then $n - \mathrm{rank}(W)$, which is $n - (d+1)$ if $W$ is full rank.
	We also claim that $F \cap \simplex \subseteq \gamma_r$.
	To see this, consider $q \in F$, but $q \not \in \gamma_r$.
	That would mean that $Wq + b = \vec 0 \iff W'q + b' = \vec 0$ (since $q \in \simplex$ satisfies the simplex constraints).
	This in turn implies $\vec 0 \in \partial L(u;q)$, implying $q \in \Gamma_u$.
	Therefore, $q \not \in \gamma_r$ would contradict $L$ indirectly eliciting $\gamma$.
\end{proof}

@@

\begin{lemma}\label{lem:feas-sub-is-a-flat}
	Suppose we have the finite elicitable property $\gamma$ and distribution $p \in \relint{\simplex}$ with $r \in \gamma(p)$.
	If $F$ is a flat containing $p$ such that $F \cap \simplex \subseteq \gamma_r$, then $F - p$ is a subspace contained in $\S_{\gamma_r}(p)$.
%  \raf{I think you want $p$ in the interior of the simplex for now}
\end{lemma}
\begin{proof}
  To spell it out, observe $F-p$ is a subspace as it is a linear shift of $F$, which is a subspace by definition of a flat.
  Now consider $v \in F - p$.
  Since $p \in \relint{\simplex}$, there is an open ball of radius $\epsilon$ in the same affine space as $\simplex$ so that for all $q \in B(p, \epsilon)$, we have $q \in \simplex$.
  In particular, take $\alpha = \epsilon / 2$, and we observe $p \pm \alpha v \in B(p, \epsilon)$, and therefore $p \pm \alpha v \in \simplex$.
  Moreover, by the assumption $v \in F - p$, we also have $p \pm \alpha v \in \gamma_r$. 
  Since level sets of elicitable properties are convex (\cite{lambert2009eliciting}) this is true for all $\alpha' \leq \alpha$.
  Therefore, we observe $v \in S_{\gamma_r}(p)$, so $F-p \subseteq S_{\gamma_r}(p)$.
%  First, if $p+v \in \gamma_r$ and $p -v \in \gamma_r$, then we have $v \in \S_{\gamma_r}(p)$ with $\epsilon_0 = 1$ as level sets of elicitable properties are convex by~\cite{lambert2009eliciting}.
%  If either $p + v$ or $p - v \not \in \gamma_r$, it must be because the term is out of the simplex by definition of $F$.
%  
%  However, if both $p + \alpha^+ v$ and $p - \alpha^- v \in \simplex$ for some $\alpha^\pm \in (0,1)$, then $v \in \S_{\gamma_r}(p)$ with $\epsilon_0 = \min(\alpha^+, \alpha^-)$.
%  As $p \in \relint{\simplex}$, there is always such an $\epsilon_0$; if there were not, then we would observe $v \not \in F - p$.
%  Therefore, we have $v \in F - p \implies v \in \S_{\gamma_r}(p)$, so $(F - p) \subseteq \S_{\gamma_r}(p)$.
\end{proof}

@@

\begin{corollary}[\cite{ramaswamy2016convex}]
	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma$.
	Take $p \in \simplex$ and $r \in \R$ such that $p \in \gamma_r$.
	\begin{equation}
	\ccdim(\ell) \geq \|p\|_0 - \dim(\S_{\gamma_r}(p)) - 1~.~
	\end{equation}
\end{corollary}

@@

\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some identifiable $\Gamma:\simplex\to\reals^d$ with $\iden(\Gamma)=d$.
  If $\lbar$ is non-constant on every non-singleton level set of $\Gamma$, then $\eliccvx(\lbar) \geq \min(d+1,n-1)$.
\end{theorem}
\begin{proof}
  Suppose for a contradiction that we have some convex-elicitable property $\hat\Gamma:\simplex\to\reals^d$ and link $\psi : \reals^d \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  The condition $\elic_\I(\Gamma)=d$ implies that $\Gamma$ is $d$-identifiable but not $(d-1)$-identifiable, and thus there must be some level set $\Gamma_r$ which is a flat of dimension $n-d-1$.
  \raft{What I mean is you can't describe $\Gamma_r$ using a flat of another dimension; revisit once we settle terminology}

>>>>

\begin{definition}[Flat]
	A \emph{flat} $F$ of $\reals^n$ is an affine subspace of $\reals^n$.
	In other words, $F\subseteq \reals^n$ can be written $F=\{x\in\reals^n : Wx + b = 0\}$ for some $W\in\reals^{d\times n}$ and $b\in\reals^d$ where $d\leq n$.
	The codimension of the flat $F$ is given by $f := \mathrm{rank}(W)$.
\end{definition}

\begin{theorem}\label{thm:cvx-flats}
	Suppose we are given a property $\gamma$ and distribution $p \in \simplex$.
	For all $r\in\gamma(p)$, if there is no flat $F$ of codimension at most $d$ containing $p$ so that $F \cap \simplex \subseteq \gamma_r$, then there is no convex surrogate loss $L : \reals^d \times \Y \to \reals$ that indirectly elicits $\gamma$.
\end{theorem}
\begin{proof}[Proof]
	We prove the contrapositive: if there is a surrogate loss $L$ indirectly eliciting $\gamma$, then for all $p \in \simplex$ and $r \in \R$ such that $r \in \gamma(p)$, there is an $d$-codimensional flat $F$ such that $F \cap \simplex \subseteq \gamma_r$ and $p \in F$.

@@

	Observe that the Minkowski sum $\oplus_y p_y \partial L(u,y) = \{p_1 x_1 + \ldots + p_n x_n : x_1 \in \partial L(u,1), \ldots, x_n \in \partial L(u,n) \}$.
	In order for $\vec 0$ to be in this $p$-weighted Minkowski sum, we then have that each choice of $x_i$ is a flat for $\partial L(u, i)$.
	Moreover, with $p$ fixed, we know there is a vector of witnesses $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$ such that $\sum_y p_y x_y = \vec 0$.
	Now we can take $F := \ker(W)$, whose rank is at most $d$.

%	Observe that $F' := \ker(W')$ is at most $(n-d)$ dimensions in $\reals^n$ and contains $p$ by construction.
%	Since we are only concerned with a flat intersected with the simplex being a subset of $\gamma_r$, we can reduce the dimension by adding an affine simplex constraint. 
%	That is, take $F := \{x \in \reals^n : W x + b = \vec 0\}$, where we have $W := [W'; \mathbbm{1}^T] \in \reals^{(d+1) \times n}$ and $b \in \reals^{d+1}$.
%	The dimension of $F$ is then $n - \mathrm{rank}(W)$, which is $n - (d+1)$ if $W$ is full rank.
	We also claim that $F \cap \simplex \subseteq \gamma_r$.
	To see this, consider $q \in F$, but $q \not \in \gamma_r$.
	That would mean that $Wq + b = \vec 0$ , which in turn implies $\vec 0 \in \partial L(u;q)$, implying $q \in \Gamma_u$.
	Therefore, $q \not \in \gamma_r$ would contradict $L$ indirectly eliciting $\gamma$.
\end{proof}

@@

\begin{lemma}\label{lem:feas-sub-is-a-flat}
	Suppose we have the finite elicitable property $\gamma$ and distribution $p \in \relint{\simplex}$ with $r \in \gamma(p)$.
	If $F$ is a flat containing $p$ such that $F \cap \simplex \subseteq \gamma_r$, then $F - p$ is a subspace contained in $\S_{\gamma_r}(p)$.
%  \raf{I think you want $p$ in the interior of the simplex for now}
\end{lemma}
\begin{proof}
  To spell it out, observe $F-p$ is a subspace as it is a linear shift of $F$, which is a linear subspace by definition of a flat and the fact that it contains $\vec 0$.
  Now consider $v \in F - p$.
  Since $p \in \relint{\simplex}$, there is an open ball of radius $\epsilon$ in the affine hull of $\simplex$ so that for all $q \in B(p, \epsilon)$, we have $q \in \simplex$.
  In particular, take $\alpha = \epsilon / 2$, and we observe $p \pm \alpha v \in B(p, \epsilon)$, and therefore $p \pm \alpha v \in \simplex$.
  Moreover, by the assumption $v \in F - p$, we also have $p \pm \alpha v \in \gamma_r$. 
  Since level sets of elicitable properties are convex (\cite{lambert2009eliciting}) this is true for all $\alpha' \leq \alpha$.
  Therefore, we observe $v \in S_{\gamma_r}(p)$, so $F-p \subseteq S_{\gamma_r}(p)$.
%  First, if $p+v \in \gamma_r$ and $p -v \in \gamma_r$, then we have $v \in \S_{\gamma_r}(p)$ with $\epsilon_0 = 1$ as level sets of elicitable properties are convex by~\cite{lambert2009eliciting}.
%  If either $p + v$ or $p - v \not \in \gamma_r$, it must be because the term is out of the simplex by definition of $F$.
%  
%  However, if both $p + \alpha^+ v$ and $p - \alpha^- v \in \simplex$ for some $\alpha^\pm \in (0,1)$, then $v \in \S_{\gamma_r}(p)$ with $\epsilon_0 = \min(\alpha^+, \alpha^-)$.
%  As $p \in \relint{\simplex}$, there is always such an $\epsilon_0$; if there were not, then we would observe $v \not \in F - p$.
%  Therefore, we have $v \in F - p \implies v \in \S_{\gamma_r}(p)$, so $(F - p) \subseteq \S_{\gamma_r}(p)$.
\end{proof}

@@

%\begin{lemma}\label{lem:p-boundary-fstar}
%	For a given property $\gamma$, fix $p \in \simplex$ and $r \in\R$ so that $r \in \gamma(p)$.
%	Consider the flat $F'$ containing $p$ of dimension at least $n-d-1$.
%	The flat $F^* = F' \cap \spn(\{e_y : y \in \supp(p)\})$ has dimension $\dim(F') - (n - \|p\|_0)$.
%\end{lemma}
%\begin{proof}
%	Consider $P := \spn(\{e_y : y \in \supp(p) \})$.
%	\begin{align*}
%	\dim(F^*) &= \dim(F') + \|p\|_0 - \dim(F' + P)
%	\end{align*}
%	The claim then holds if $n = \dim(F' + P)$.
%	\jessie{This is where I'm stuck??}
%\end{proof}

\begin{lemma}\label{lem:p-boundary-fsd}
	For any $p \in \simplex$ and $r$ such that $p \in \gamma_r$, define $\gamma'$ to be $\gamma$ restricted to the simplex $\Delta_{\supp(p)}$.
	Consider the ambient space of both $\S_{\gamma_r}(p)$ and $\S_{\gamma'_r}(p)$ to be $\reals^\Y$.
	Then $\S_{\gamma_r}(p) = \S_{\gamma'_r}(p)$.
\end{lemma}
\begin{proof}
	We trivially have $\S_{\gamma_r}(p) \supseteq \S_{\gamma'_r}(p)$ since the latter is simply a projection down to an affine subspace of the former.

	Now to see $\S_{\gamma_r}(p) \subseteq \S_{\gamma'_r}(p)$, take some $v \not \in \S_{\gamma'_r}(p)$.
	Then for all $\epsilon > 0$, we have either $p \pm \epsilon v \not \in \gamma'_r$.
	Moreover, this must be because $p \pm \epsilon v$ leaves the simplex $\Delta_{\supp(p)}$, as we have $p$ on the relative interior of that such simplex.
	While $p \pm \epsilon v$ may be in $\simplex$, we must then have $p \mp \epsilon v$ outside the simplex.
	As $\gamma_r \subseteq \simplex$, we have $v \not \in \S_{\gamma_r}(p)$.

	%	For simplicity, consider $\epsilon = \epsilon_0 / 2$.
%	Consider $P := \spn(\{e_y : y \in \supp(p) \})$.
%	\begin{align*}
%	\dim(\S'_{\gamma_r}(p)) &= \dim(\S_{\gamma_r}(p)) + \dim(P) - \dim(\S_{\gamma_r}(p) + P)
%	\end{align*}
%	If $\dim(P) = \dim(\S_{\gamma_r}(p) + P)$, the claim holds.
%	Be definition of $P$, we know $\dim(P) = \|p\|_0$, so we can show $\dim(\S_{\gamma_r}(p) + P) = \|p\|_0$.		
%	One way to do this is to show $\S_{\gamma_r}(p) \subseteq P$.
%	
%	Suppose there was some $v \not\in P$ such that $v \in \S_{\gamma_r}(p)$.
%	That means there is some $i$ so that $i \not \in \supp(p)$ (i.e. $p_i = 0$) but $v_i \neq 0$. \jessiet{Check this line.}
%	However, if $v \in \S_{\gamma_r}(p)$, then there is some $\epsilon_0 > 0$ so that $p + \epsilon v$ and $p - \epsilon v \in \gamma_r \subseteq \simplex$ for all $\epsilon \in (0, \epsilon_0)$.
%	For simplicity, consider $\epsilon = \epsilon_0 / 2$.
%	Since $p_i = 0$, this gives $(p+\epsilon v)_i = (\epsilon v)_i$, and likewise for $(p - \epsilon v)_i$.
%	As neither $\epsilon$ nor $v_i$ are $0$, one of these terms must be negative, and therefore not in the simplex.
%	As $\gamma_r \subseteq \simplex$, we then have $v \not \in \S_{\gamma_r}(p)$.
%	Therefore, $\S_{\gamma_r}(p) \subseteq P$, and $\dim(\S_{\gamma_r}(p) + P) = \dim(P) = \|p\|_0$.
\end{proof}

\begin{corollary}[\cite{ramaswamy2016convex}]
	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma$.
	Take $p \in \simplex$ and $r \in \R$ such that $p \in \gamma_r$.
	\begin{equation}
	\ccdim(\ell) \geq \|p\|_0 - \dim(\S_{\gamma_r}(p)) - 1~.~
	\end{equation}
\end{corollary}
\begin{proof}
	Let $L : \reals^d \times \Y \to \reals$ be a calibrated surrogate for $\ell$.
	First, for intuition, consider $p \in \relint{\simplex}$. 
	Theorem~\ref{thm:cvx-flats} implies that there exists a flat $F'$ of affine dimension at least $n-d-1$.
	Lemma~\ref{lem:feas-sub-is-a-flat} then says that $\dim(\S_{\gamma_r}(p)) \geq \dim(F') \geq n-d-1 \implies d \geq n - \dim(\S_{\gamma_r}(p)) - 1$.

	Now, consider the distribution $p_0$ over $\Y' := \supp(p)$ where each component of $p_0$ is equal to the corresponding entry of $p$.

	If we can show that $L$ is calibrated with respect to $\ell' := \R \times \Y' \to \reals$ with $\ell(r,y) = \ell'(r,y)$ for all $r \in \R$ and $y \in \Y'$, then we observe the existence of a flat $F^*$ of codimension (in $\reals^{\Y'}$) at most $d$ so that $\dim(F') = \|p\|_0 - d - 1$ by Theorem~\ref{thm:cvx-flats}.
	Then we can use Lemma~\ref{lem:p-boundary-fsd} to observe $\dim(\S_{\gamma'_r}(p_0)) = \dim(\S_{\gamma_r}(p))$, and thus the result holds.

	Now to see that $L$ is calibrated with respect to $\ell'$, consider that $\gamma(p) = \gamma'(p)$ for all $p \in \Delta^{\Y'}$, and thus the bound in Equation~\eqref{eq:calibration} is true for all $p \in \Delta_{\Y'}$ as the expected losses are equal for all $y \in \Y'$.
%	Now, if $p \not \in \relint{\simplex}$, we can apply Lemma~\ref{lem:p-boundary-fstar} to observe the existence of an $F^*$ with dimension $\dim(F') - (n - \|p\|_0)$.
%	Moreover, since projecting to $(\{e_y : y \in \supp(p)\})$ does not change the feasible subspace dimension, so we observe the result.
\end{proof}

@@

\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some identifiable $\Gamma:\simplex\to\reals^d$ with $\iden(\Gamma)=d$.
  If $\lbar$ is non-constant on every non-singleton level set of $\Gamma$, then $\eliccvx(\lbar) \geq \min(d+1,n-1)$.
\end{theorem}
\begin{proof}
  Suppose for a contradiction that we have some convex-elicitable property $\hat\Gamma:\simplex\to\reals^d$ and link $\psi : \reals^d \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  The condition $\iden(\Gamma)=d$ implies that $\Gamma$ is $d$-identifiable but not $(d-1)$-identifiable, and thus there must be some level set $\Gamma_r$ which is a flat of dimension $n-d-1$.
  \raft{What I mean is you can't describe $\Gamma_r$ using a flat of another dimension; revisit once we settle terminology}
